### 2011-02-27 paper/paper.tex

\documentclass{article}
\newcommand{\osn}{\oldstylenums}
\title{TITLE HERE}
\author{Alec Story and Thomas Levine\\\small{avs\osn{38} and tkl\osn{22}}}

\begin{document}
\maketitle
\section*{Sentence Generation}

	\subsection*{Unsmoothed}

		\subsubsection*{Unigrams}

			\begin{quote}

			Bring in Take made the pass a Caesar in gold well leave \#

			Exeunt . hope , and Milan your A Will thy have Come the . .
			advantage , together your either to '\#

			with . ratified who of are in my year to natural I have the do to
			Master Turkish there wherein \#

			\end{quote}

		\subsubsection*{Bigrams}

			\begin{quote}

			Enter LYSANDER , and you lie dull actor on my speech , an ordinary
			pitch , Loyal and full of our tongues , see , That which physic to
			my horse for her youth and sealed bags Of base earth ; Though twenty
			several mistress will you have done To have found . \#

			Silence , who . \#

			The king , go away ! \#

			An heir ; for the likeness of eternal in a raven black , the sons ,
			whisper me have all those that thou wert not me welcome . \#

			\end{quote}

		\subsubsection*{Trigrams}

			\begin{quote}

			Re-enter BIONDELLO , LUCENTIO , and goodness never fearful . \#

			Gracious , my lad . \#

			How - traitor ? \#

			The Greeks upon advice , Hath alter'd that good wisdom Whereof I know
			by their christen names , Unless thou tell her so . \#

			\end{quote}

	\subsection*{Smoothed}

\end{document}

### 2011-02-27 experiment.py
run an experiment

@@ -1,0 +1,26 @@
+#!/usr/bin/env python
+import ngram
+import filters
+import csv
+import random
+
+ngram.good_turing =lambda x:x
+
+def trial(train,n,p='ap'):
+	"""n is n-gram size. p is whether it's log, arbitrary precision or float"""
+	#Start timer
+	probs=ngram.probabilities(ngram.good_turing(ngram.ngram(n,filters.unk(filters.shakespeare(train)))))
+	sentence=ngram.make_sentence(probs)
+	pp=ngram.perplexity(probs,sentence)
+	#End timer
+	time=42
+	return [n,p,pp,time,sentence]
+
+def sentence_generation(train,filename,nmax,reps):
+	out=csv.writer(open(filename, 'wb'), delimiter='|', quotechar='&', quoting=csv.QUOTE_NONE)
+	for i in range(0,nmax*reps):
+		n=random.randint(1,nmax)
+		out.writerow(trial(train,n))
+
+def main():
+	sentence_generation('Shakespeare/Train.txt','hi.csv',3,30)

### 2011-02-27 ngram.py
main function for profiling

@@ -1,12 +1,11 @@
 #!/usr/bin/env python
-import gmpy
-import mpmath as mp
 from fractions import Fraction
 from math import log,e
-import math
-import random
-from random import normalvariate #Just for generating numbers for testing functions
 import filters
+import gmpy
+import math
+import mpmath as mp
+import random
 
 use_fractions = True
 
@@ -126,7 +125,6 @@ def make_sentence(probs):
 		word_list = []
 	word_buffer = list(word_list)
 	while(True):
-		# TODO(astory): unk
 		prob_list = [(x) for x in probs[len(word_buffer)+1].items()\
 				if x[0][0:-1] == tuple(word_buffer)]
 		ngram = choose_prob(prob_list)
@@ -150,5 +148,13 @@ def perplexity_log(W):
 	PP=(-1/N)*p_log(W)
 	return PP
 
+def main():
+	words = filters.shakespeare('Shakespeare/Train.txt')
+	probs = ngram(3, words)
+	random.jumpahead(3474)
+	print make_sentence(probs)
+
+if __name__ == "__main__":
+	main()
 
 #ngram.make_sentence(ngram.probabilities(ngram.good_turing(ngram.ngram(3,filters.unk(filters.shakespeare('Shakespeare/short.txt'))))))

### 2011-02-27 experiment.py
faster experiment but broken

@@ -6,21 +6,28 @@ import random
 
 ngram.good_turing =lambda x:x
 
-def trial(train,n,p='ap'):
+def trial(train,n,probs,p='ap'):
 	"""n is n-gram size. p is whether it's log, arbitrary precision or float"""
 	#Start timer
-	probs=ngram.probabilities(ngram.good_turing(ngram.ngram(n,filters.unk(filters.shakespeare(train)))))
-	sentence=ngram.make_sentence(probs)
-	pp=ngram.perplexity(probs,sentence)
+	#import code
+	#code.interact(local=locals())
+	sentence=ngram.make_sentence(probs[0:(n+1)])
+	pp=ngram.perplexity(probs[0:(n+1)],sentence)
 	#End timer
 	time=42
 	return [n,p,pp,time,sentence]
 
-def sentence_generation(train,filename,nmax,reps):
+def sentence_generation(train,filename,nmax,reps,probs):
 	out=csv.writer(open(filename, 'wb'), delimiter='|', quotechar='&', quoting=csv.QUOTE_NONE)
 	for i in range(0,nmax*reps):
 		n=random.randint(1,nmax)
-		out.writerow(trial(train,n))
+		out.writerow(trial(train,n,probs))
 
 def main():
-	sentence_generation('Shakespeare/Train.txt','hi.csv',3,30)
+	nmax=3
+	reps=3
+	train='Shakespeare/short.txt'
+	out='results.csv'
+	#TODO(tom) This isn't working. I'll fix it later.
+	probs=ngram.probabilities(ngram.good_turing(ngram.ngram(nmax,filters.unk(filters.shakespeare(train)))))
+	sentence_generation(train,out,nmax,reps,probs)

### 2011-02-27 experiment.py
War and Peace works now

@@ -26,7 +26,8 @@ def sentence_generation(train,filename,nmax,reps,probs):
 def main():
 	nmax=3
 	reps=3
-	train='Shakespeare/short.txt'
+	#train='Shakespeare/short.txt'
+	train='War and Peace/short.txt'
 	out='results.csv'
 	#TODO(tom) This isn't working. I'll fix it later.
 	probs=ngram.probabilities(ngram.good_turing(ngram.ngram(nmax,filters.unk(filters.shakespeare(train)))))

### 2011-02-28 ngram.py
sentence generation accounts for logs

@@ -140,6 +140,8 @@ def probabilities(ngrams):
 	
 
 def choose_prob(l):
+	if not use_fractions:
+		l = [(t, exp(p)) for (t,p) in l]
 	total_prob = 0
 	for (t,v) in l:
 		total_prob += v

### 2011-02-28 ngram.py
removed code.interact()...

@@ -61,8 +61,6 @@ def good_turing(ngrams):
 	#Put this in a dictionary in the format of ngrams
 	ngrams_smoothed=[{}]
 	n_c=get_n_c(ngrams)
-	import code
-	code.interact(local = locals())
 	for n in range(1,len(ngrams)):
 		ngrams_smoothed.append({})
 		for gram in ngrams[n].keys():

### 2011-02-28 ngram.py
better progress indicator for make sentences

@@ -165,6 +165,7 @@ def make_sentence(probs):
 		word_buffer.append(w)
 		if len(word_buffer) == n:
 			word_buffer.pop(0)
+		print " ".join(word_list)
 		if w == filters.SOS:
 			break
 	return word_list

### 2011-02-28 experiment.py
blarg

@@ -6,22 +6,25 @@ import random
 
 ngram.good_turing =lambda x:x
 
-def trial(train,n,probs,p='ap'):
+def trial(train,n,p,probs):
 	"""n is n-gram size. p is whether it's log, arbitrary precision or float"""
 	#Start timer
-	#import code
-	#code.interact(local=locals())
 	sentence=ngram.make_sentence(probs[0:(n+1)])
 	pp=ngram.perplexity(probs[0:(n+1)],sentence)
 	#End timer
-	time=42
+	time='NA'
 	return [n,p,pp,time,sentence]
 
 def sentence_generation(train,filename,nmax,reps,probs):
 	out=csv.writer(open(filename, 'wb'), delimiter='|', quotechar='&', quoting=csv.QUOTE_NONE)
+	out.writerow(['n','use fractions','perplexity','time','sentence'])
 	for i in range(0,nmax*reps):
 		n=random.randint(1,nmax)
-		out.writerow(trial(train,n,probs))
+		#ngram.use_fractions=random.randint(0,1)
+		ngram.use_fractions=True
+		out.writerow(trial(train,n,ngram.use_fractions,probs))
+		ngram.use_fractions=False
+		out.writerow(trial(train,n,ngram.use_fractions,probs))
 
 def main():
 	nmax=3

### 2011-02-28 ngram.py

@@ -3,7 +3,7 @@ import copy
 import gmpy
 import mpmath as mp
 from fractions import Fraction
-from math import log,e
+from math import log,e,exp
 import filters
 import gmpy
 import math
@@ -61,8 +61,6 @@ def good_turing(ngrams):
 	#Put this in a dictionary in the format of ngrams
 	ngrams_smoothed=[{}]
 	n_c=get_n_c(ngrams)
-	import code
-	code.interact(local = locals())
 	for n in range(1,len(ngrams)):
 		ngrams_smoothed.append({})
 		for gram in ngrams[n].keys():
@@ -110,6 +108,7 @@ def perplexity(probs, words):
 					break
 				else:
 					temp_words.pop(0)
+	print use_fractions
 	if use_fractions:
 		return mp.power((mp.mpf(prob.denom())/mp.mpf(prob.numer())), 1.0/len(words))
 	else:
@@ -140,17 +139,17 @@ def probabilities(ngrams):
 	
 
 def choose_prob(l):
-	if not use_fractions:
+	if l[0][1]<=0:
 		l = [(t, exp(p)) for (t,p) in l]
 	total_prob = 0
 	for (t,v) in l:
 		total_prob += v
 	n = random.uniform(0,total_prob)
 	for item, weight in l:
-		if n < weight:
+		if n <= weight:
 			return item
 		n -= weight
-
+	
 def make_sentence(probs):
 	n = len(probs) - 1
 	if n > 1:
@@ -162,6 +161,8 @@ def make_sentence(probs):
 		prob_list = [(x) for x in probs[len(word_buffer)+1].items()\
 				if x[0][0:-1] == tuple(word_buffer)]
 		ngram = choose_prob(prob_list)
+		if ngram==None:
+			print 'n-gram is none'
 		w = ngram[-1]
 		word_list.append(w)
 		word_buffer.append(w)

### 2011-02-28 filters.py
Lots of changes

@@ -47,8 +47,8 @@ def shakespeare(filename):
 					words.append(word[-1]) #punctuation
 				elif word[0] in string.punctuation:
 					#beginning-of-word mid-sentence punctuation
-					words.append(word[-1]) #punctuation
-					words.append(word[0:-1]) #word
+					words.append(word[0]) #punctuation
+					words.append(word[1:]) #word
 				#This currently does not handle middle-of-word punctuation like someone else's program would.
 				else:
 					words.append(word)

### 2011-02-28 ngram.py

@@ -2,12 +2,10 @@
 import copy
 import gmpy
 import mpmath as mp
-from fractions import Fraction
 from math import log,e
 import filters
 import gmpy
 import math
-import mpmath as mp
 import random
 
 use_fractions = True
@@ -81,7 +79,7 @@ def good_turing(ngrams):
 def perplexity(probs, words):
 	n = len(probs) - 1
 	if use_fractions:
-		prob = Fraction(1,1)
+		prob = gmpy.mpq(1,1)
 	else:
 		prob = 0
 	word_list = []
@@ -169,17 +167,6 @@ def make_sentence(probs):
 		if w == filters.SOS:
 			break
 	return word_list
-
-#Small things that we need to add at some point
-def p_log(W):
-	return 0.00023
-
-def perplexity_log(W):
-	#W = w_1 w_2 ... w_N
-	#PP (W) = P (w1 w2 .. wN) ^ -1/N
-	N=len(W)
-	PP=(-1/N)*p_log(W)
-	return PP
 
 def main():
 	words = filters.shakespeare('Shakespeare/Train.txt')

### 2011-02-28 paper/paper.tex

@@
 
+	For sentence generation, we did not include our unknown processing, and do
+	not convert the first occurrence of each word to an unknown token.
+
+	Each subsection contains two examples from the Shakespeare corpus and two
+	from War and Peace.  These corpora were chosen because they were determined
+	at rough glance to be the easiest to parse and had the fewest problematic
+
 	\subsection*{Unsmoothed}
@@
 
-			Bring in Take made the pass a Caesar in gold well leave \#
+			Bring in Take made the pass a Caesar in gold well leave
 
 			Exeunt . hope , and Milan your A Will thy have Come the . .
-			advantage , together your either to '\#
+			advantage , together your either to '
 
-			with . ratified who of are in my year to natural I have the do to
-			Master Turkish there wherein \#
+			He fell other along " therefore , and had about He chief shorten now
+			him life . something
+
+			I had transference shrugged man with you The lit . to lift the her
+			had he . and have the hand individuals dragging the , place been to
+			Natasha in inside . revelry hide nation room beaming eager all of
+			the the , remounts to
 
@@
 			my horse for her youth and sealed bags Of base earth ; Though twenty
-			several mistress will you have done To have found . \#
+			several mistress will you have done To have found .
 
-			Silence , who . \#
+			Silence , who .
 
-			The king , go away ! \#
+			The horses ' familiar to the army , envy and looked questioningly :
+			"You won't be thus depriving him of the estate with his will send me
+			forty thousand men in choosing her governess kept saying good-by .
 
-			An heir ; for the likeness of eternal in a raven black , the sons ,
-			whisper me have all those that thou wert not me welcome . \#
+			I have begun to Tushin and yet further battles , the man in the
+			defense of that temple , they were standing beside whom you seen in
+			the battle of dry branches of meeting it has brought the smile Berg
+			with a mocking , and even impossible to be long paced angrily .
 
@@
 
-			Re-enter BIONDELLO , LUCENTIO , and goodness never fearful . \#
-
-			Gracious , my lad . \#
-
-			How - traitor ? \#
+			How - traitor ?
 
 			The Greeks upon advice , Hath alter'd that good wisdom Whereof I know
-			by their christen names , Unless thou tell her so . \#
+			by their christen names , Unless thou tell her so .
+
+			No , but it can't be helped It happens to the first to arrive ,
+			Princess , I think .
+
+			Rostopchin , coming to Petersburg , look out at night his feet .
 
@@
 
+		\subsubsection*{Unigram}
+
+			\begin{quote}
+
+			th ! come Lear things of the I
+
+			me . I
+
+			they its intimate just such But There the them said shout
+			irresistible always heard none with and of dinner had To
+
+			day good-natured Secondly same , joyful on talking the appeared
+			advantage . unreasoning men by
+
+			\end{quote}
+
+		\subsubsection*{Bigram}
+
+			\begin{quote}
+
+			Hector .
+
+			I arrest thee here , which time , that his fault , and blows .
+
+			But my wife should have nobody and ask the enemy was too .
+
+			Say it's farther .
+
+			\end{quote}
+
+		\subsubsection*{Trigram}
+
+			\begin{quote}
+
+			She would have made fair work !
+
+			Come , my lord , I'll bring you to please his Majesty , Herod of
+			Jewry dare not say he is much abus'd with tears thou keep'st command
+			.
+
+			All that he had no news from the chase .
+
+			If I decline the honor of the bushes .
+
+			\end{quote}
+
+	\section*{Arbitrary Precision}
+
+		One of the two extensions we chose was to implement the probability and
+		perplexity calculations using arbitrary precision arithmetic (we also
+		implemented it using logarithmic arithmetic).  This can lead to slower
+		processing, but in our profiling, most of the time was spent in
+		dictionary manipulation rather than mathematics, and guarantees that no
+		information is lost.
+
+		We used python's +gmpy+ package to represent the probabilities as
+		arbitrary rational numbers (python's integers are already
+		arbitrary-precision), and +mpmath+ to perform the final exponentiation
+		to calculate the perplexity (the $n$-th root).
+
+		Our hypothesis was that the arbitrary precision math would reveal a
+		small amount of error in computing the perplexity logarithmically.
+
+	\section*{Arbitrary $n$-grams}
+
+		Because the computations were general enough, we were able to easily
+		abstract the code to produce count, smoothed count, and probability
+		tables for arbitrary $n$-grams.  We also compute all of the smaller
+		$n$-grams, which are required for generating or analyzing the beginning
+		of a sequence of words, and for other computations.
 \end{document}

### 2011-02-28 experiment.py
output data, read to R

@@ -13,7 +13,7 @@ def trial(train,n,p,probs):
 	pp=ngram.perplexity(probs[0:(n+1)],sentence)
 	#End timer
 	time='NA'
-	return [n,p,pp,time,sentence]
+	return [n,p,pp,time,' '.join(sentence)]
 
 def sentence_generation(train,filename,nmax,reps,probs):
 	out=csv.writer(open(filename, 'wb'), delimiter='|', quotechar='&', quoting=csv.QUOTE_NONE)

### 2011-02-28 paper/analysis.R

@@ -1,0 +1,1 @@
+results=read.csv('results.csv',sep='|',colClasses=c('integer','factor','numeric','numeric','character'))

### 2011-02-28 paper/analysis.R
general results importing

@@ -1,1 +1,1 @@
-results=read.csv('results.csv',sep='|',colClasses=c('integer','factor','numeric','numeric','character'))
+get_results=function(file) {read.csv(file,sep='|',colClasses=c('integer','logical','numeric','numeric','character'))}

### 2011-02-28 experiment.py
fuller experiment

@@ -18,20 +18,25 @@ def trial(train,n,p,probs):
 def sentence_generation(train,filename,nmax,reps,probs):
 	out=csv.writer(open(filename, 'wb'), delimiter='|', quotechar='&', quoting=csv.QUOTE_NONE)
 	out.writerow(['n','use fractions','perplexity','time','sentence'])
-	for i in range(0,nmax*reps):
-		n=random.randint(1,nmax)
-		#ngram.use_fractions=random.randint(0,1)
-		ngram.use_fractions=True
-		out.writerow(trial(train,n,ngram.use_fractions,probs))
-		ngram.use_fractions=False
-		out.writerow(trial(train,n,ngram.use_fractions,probs))
+	for i in range(0,reps):
+		for n in range(1,nmax):
+			try:
+				#ngram.use_fractions=random.randint(0,1)
+				ngram.use_fractions=True
+				out.writerow(trial(train,n,ngram.use_fractions,probs))
+				ngram.use_fractions=False
+				out.writerow(trial(train,n,ngram.use_fractions,probs))
+			except(AttributeError):
+				pass
 
-def main():
-	nmax=3
-	reps=3
+def run(train,nmax,out):
+	reps=30
 	#train='Shakespeare/short.txt'
-	train='War and Peace/short.txt'
-	out='results.csv'
 	#TODO(tom) This isn't working. I'll fix it later.
 	probs=ngram.probabilities(ngram.good_turing(ngram.ngram(nmax,filters.unk(filters.shakespeare(train)))))
 	sentence_generation(train,out,nmax,reps,probs)
+
+def main():
+	nmax=5
+	run('War and Peace/Train.txt',nmax,'War and Peace results.csv')
+	run('Shakespeare/Train.txt',nmax,'Shakespeare results.csv')

### 2011-02-28 paper/analysis.R
plot of perplexity by n and use_fractions

@@ -1,1 +1,7 @@
 get_results=function(file) {read.csv(file,sep='|',colClasses=c('integer','logical','numeric','numeric','character'))}
+
+#results.shakespeare=get_results('Shakespeare results.csv')
+results.tolstoy=get_results('War and Peace results.csv')
+
+foo=results.tolstoy
+par(mfrow=c(1,2));for (i in c(0,1)){plot((perplexity)~n,data=subset(foo,use.fractions==i),main=paste('use fractions =',i))}

### 2011-02-28 experiment.py
off by one in experiment nmax

@@ -19,7 +19,7 @@ def sentence_generation(train,filename,nmax,reps,probs):
 	out=csv.writer(open(filename, 'wb'), delimiter='|', quotechar='&', quoting=csv.QUOTE_NONE)
 	out.writerow(['n','use fractions','perplexity','time','sentence'])
 	for i in range(0,reps):
-		for n in range(1,nmax):
+		for n in range(1,nmax+1):
 			try:
 				#ngram.use_fractions=random.randint(0,1)
 				ngram.use_fractions=True

### 2011-02-28 paper/paper.tex
Changes, fixes, etc.

@@
 \newcommand{\osn}{\oldstylenums}
-\title{TITLE HERE}
+\title{Arbitrary Precision and Arbitrary $n$-grams}
 \author{Alec Story and Thomas Levine\\\small{avs\osn{38} and tkl\osn{22}}}
@@
 	from War and Peace.  These corpora were chosen because they were determined
-	at rough glance to be the easiest to parse and had the fewest problematic
+	at rough glance to be the easiest to parse and had the fewest problematic.
 
@@
 
+		Trigrams lead to more comprehensible sentences, which is to be expected,
+		and bigrams are barely useful.  The unigrams also seemed to produce
+		longer sentences; it's possible that trigrams and bigrams lead to a
+		progression from words that are likely at the beginning of sentences to
+		words that are likely at the end, and then to end the sentence, despite
+		sentence ends being fairly infrequent.  Additionally, \textit{War and
+		Peace} seems to produce longer sentences (particularly when observed
+		over more iterations than can be reproduced comfortably here), probably
+		because the sentences in the original are longer, so sentence ends are
+		rarer.
+
+		Our $n$-gram generator only treated sentence endings by adding a single
+		symbol, `\#', and did not break the input into individual sentences, nor
+		do we break the perplexity input into sentences in the same manner.
+
 	\section*{Arbitrary Precision}
@@
 		processing, but in our profiling, most of the time was spent in
-		dictionary manipulation rather than mathematics, and guarantees that no
-		information is lost.
+		dictionary manipulation rather than mathematics.  This approach
+		guarantees that there is no data loss until we move into processing
+		floats during exponentiation.
 
@@
 		arbitrary-precision), and +mpmath+ to perform the final exponentiation
-		to calculate the perplexity (the $n$-th root).
+		to calculate the perplexity (the $n$-th root).  The libraries could only
+		compute non-integer powers of integers or floats, and integer powers
+		were rounded to the nearest integer, and therefore, inaccurate, so
+		(very high precision) floats provided the best results.  This final
+		stage of computation of perplexity was the only point where imprecise
+		mathematics was used.
 
@@
 		of a sequence of words, and for other computations.
+
+		Large $n$-grams (particularly $n > 4$) tend to result in whole
+		Shakespeare sentence fragments.  For example, ``O , my offence is rank ,
+		it smells to heaven ; It hath the primal eldest curse upon't , A
+		brother's murther !'' was generated by a hexagram, and is a direct quote
+		from Hamlet, act \osn{3} scene \osn{3}.  This is a well-documented
+		phenomenon, and is due to the fact that it's quite possible that the
+		preceding $n-1$-gram appears only once in the entire corpus, so the
+		sentence generator has no choice but to choose the next word in the
+		quotation.
+
 \end{document}

### 2011-02-28 filters.py

@@ -24,34 +24,37 @@ def unk(words):
 			words[i]=UNK
 	return words
 
-def shakespeare(filename):
-	"""Processes input assuming it is shakespeare-like"""
+def strip_punct(word):
+	if word == '' or None:
+		l = []
+	elif word[-1] in eos :
+		#end of sentence punctuation
+		l = strip_punct(word[0:-1])
+		l.append(word[0:-1])
+		l.append(word[-1])
+		l.append(SOS)
+	elif word[-1] in string.punctuation:
+		#end-of-word mid-sentence punctuation
+		l = strip_punct(word[0:-1])
+		l.append(word[-1])
+	elif word[0] in string.punctuation:
+		#beginning-of-word mid-sentence punctuation
+		l = [word[0]]
+		l.extend(strip_punct(word[1:]))
+	else:
+		l = [word]
+	return l
+
+def shakespeare(filehandle):
 	words = [SOS]
-	f = open(filename)
+	f = filehandle
 	for line in f:
 		# if the line is empty, or is just something like "SECOND LORD", the
 		# sentence is over, and we should add a new one.
 		if valid_line_regex.search(line):
 			line_words = whitespace.split(line)
 			for word in line_words:
-				if word == '':
-					pass
-				elif word[-1] in eos :
-					#end of sentence punctuation
-					words.append(word[0:-1]) #word
-					words.append(word[-1]) #eos punctuation
-					words.append(SOS) #sos marker
-				elif word[-1] in string.punctuation:
-					#end-of-word mid-sentence punctuation
-					words.append(word[0:-1]) #word
-					words.append(word[-1]) #punctuation
-				elif word[0] in string.punctuation:
-					#beginning-of-word mid-sentence punctuation
-					words.append(word[0]) #punctuation
-					words.append(word[1:]) #word
-				#This currently does not handle middle-of-word punctuation like someone else's program would.
-				else:
-					words.append(word)
+				words.extend(strip_punct(word))
 		else:
 			if not words[-1] == SOS:
 				words.append(SOS)

### 2011-02-28 ngram.py

@@ -1,4 +1,5 @@
 #!/usr/bin/env python
+import argparse as ap
 import copy
 import gmpy
 import mpmath as mp
@@ -7,8 +8,6 @@ import filters
 import gmpy
 import math
 import random
-
-use_fractions = True
 
 #K for good-turing
 K=5
@@ -77,13 +76,19 @@ def good_turing(ngrams):
 	return ngrams_smoothed
 
 def perplexity(probs, words):
+	global use_fractions
+	print "perplexity use_fractions: %s" % use_fractions
 	n = len(probs) - 1
 	if use_fractions:
 		prob = gmpy.mpq(1,1)
 	else:
 		prob = 0
 	word_list = []
+	i = 0
 	for w in words:
+		i += 1
+		if i % 10000 == 0:
+			print "processing word %d" % i
 		word_list.append(w)
 		if len(word_list) > n:
 			word_list.pop(0)
@@ -106,13 +111,14 @@ def perplexity(probs, words):
 					break
 				else:
 					temp_words.pop(0)
-	print use_fractions
 	if use_fractions:
 		return mp.power((mp.mpf(prob.denom())/mp.mpf(prob.numer())), 1.0/len(words))
 	else:
 		return math.exp(-prob/len(words))
 
 def probabilities(ngrams):
+	global use_fractions
+	print "probabilities use_fractions: %s" % use_fractions
 	totals = [dict_sum(x) for x in ngrams]
 	probabilities = []
 
@@ -172,10 +178,52 @@ def make_sentence(probs):
 	return word_list
 
 def main():
-	words = filters.shakespeare('Shakespeare/Train.txt')
-	probs = ngram(3, words)
-	random.jumpahead(3474)
-	print make_sentence(probs)
+	parser = ap.ArgumentParser(description='Play with some ngrams')
+	parser.add_argument('-n', '--n-gram', metavar='N', type=int,
+		dest='n', action='store', default=3,
+		help='n-gram to compute')
+	parser.add_argument('-t', '--train', metavar='FILE', type=file,
+		dest='training', action='store', default='Shakespeare/Train.txt',
+		help='Train ngrams from this file')
+	parser.add_argument('-p', '--perplexity', metavar='FILE', type=file,
+		dest='perplexity', action='store', default=None, help='measure'+
+		' perplexity against this file')
+	parser.add_argument('-l', '--logs', dest='use_logs', action='store_true',
+		help='compute using logs, default is arbitrary-precision')
[…]

### 2011-02-28 experiment.py
changes

@@ -21,11 +21,10 @@ def sentence_generation(train,filename,nmax,reps,probs):
 	for i in range(0,reps):
 		for n in range(1,nmax+1):
 			try:
-				#ngram.use_fractions=random.randint(0,1)
-				ngram.use_fractions=True
-				out.writerow(trial(train,n,ngram.use_fractions,probs))
-				ngram.use_fractions=False
-				out.writerow(trial(train,n,ngram.use_fractions,probs))
+				ngram.set_fractions(True)
+				out.writerow(trial(train,n,True,probs))
+				ngram.set_fractions(False)
+				out.writerow(trial(train,n,False,probs))
 			except(AttributeError):
 				pass
 
@@ -37,6 +36,10 @@ def run(train,nmax,out):
 	sentence_generation(train,out,nmax,reps,probs)
 
 def main():
+	ngram.set_fractions(True)
 	nmax=5
 	run('War and Peace/Train.txt',nmax,'War and Peace results.csv')
 	run('Shakespeare/Train.txt',nmax,'Shakespeare results.csv')
+
+if __name__ == '__main__':
+	main()

### 2011-02-28 ngram.py

@@ -12,6 +12,10 @@ import random
 #K for good-turing
 K=5
 
+def set_fractions(fractions):
+	global use_fractions
+	use_fractions = fractions
+
 def dict_sum(d):
 	s = 0
 	for k in d.keys():

### 2011-02-28 ngram.py
return unk on empty probability list

@@ -147,6 +147,8 @@ def probabilities(ngrams):
 	
 
 def choose_prob(l):
+	if l = []
+		return filters.UNK
 	if l[0][1]<=0:
 		l = [(t, exp(p)) for (t,p) in l]
 	total_prob = 0

### 2011-02-28 filters.py
fix to strip punct

@@ -30,7 +30,6 @@ def strip_punct(word):
 	elif word[-1] in eos :
 		#end of sentence punctuation
 		l = strip_punct(word[0:-1])
-		l.append(word[0:-1])
 		l.append(word[-1])
 		l.append(SOS)
 	elif word[-1] in string.punctuation:

### 2011-02-28 paper/analysis.R
plots of perplexity of generated sentences

@@ -4,4 +4,19 @@ get_results=function(file) {read.csv(file,sep='|',colClasses=c('integer','logica
 results.tolstoy=get_results('War and Peace results.csv')
 
 foo=results.tolstoy
-par(mfrow=c(1,2));for (i in c(0,1)){plot((perplexity)~n,data=subset(foo,use.fractions==i),main=paste('use fractions =',i))}
+foo.ag=aggregate(foo$perplexity,list(use.fractions=foo$use.fractions,n=foo$n),mean)
+
+
+#par(mfrow=c(1,2));for (i in c(0,1)){plot((perplexity)~n,data=subset(foo,use.fractions==i),main=paste('use fractions =',i))}
+
+plot(perplexity~n,data=foo,type='n',axes=F,
+	main='Perplexity of generated sentences as a function of n-gram length and probability computation method',
+	xlab='n-gram length',
+	ylab='Perplexity'
+)
+axis(1,at=1:max(foo$n))
+axis(2)
+for (i in c(0,1)){
+	points(perplexity~n,data=subset(foo,use.fractions==i),pch=21,col=NA,bg=1+i)
+	lines(x~n,data=subset(foo.ag,use.fractions==i),col=1+i)
+}

### 2011-02-28 experiment.py
stuff

@@ -16,7 +16,8 @@ def trial(train,n,p,probs):
 	return [n,p,pp,time,' '.join(sentence)]
 
 def sentence_generation(train,filename,nmax,reps,probs):
-	out=csv.writer(open(filename, 'wb'), delimiter='|', quotechar='&', quoting=csv.QUOTE_NONE)
+	out_fh = open(filename, 'wb')
+	out=csv.writer(out_fh, delimiter='|', quotechar='&', quoting=csv.QUOTE_NONE)
 	out.writerow(['n','use fractions','perplexity','time','sentence'])
 	for i in range(0,reps):
 		for n in range(1,nmax+1):
@@ -25,14 +26,16 @@ def sentence_generation(train,filename,nmax,reps,probs):
 				out.writerow(trial(train,n,True,probs))
 				ngram.set_fractions(False)
 				out.writerow(trial(train,n,False,probs))
+				out_fh.flush()
 			except(AttributeError):
 				pass
 
 def run(train,nmax,out):
-	reps=30
+	reps=2
 	#train='Shakespeare/short.txt'
 	#TODO(tom) This isn't working. I'll fix it later.
-	probs=ngram.probabilities(ngram.good_turing(ngram.ngram(nmax,filters.unk(filters.shakespeare(train)))))
+	fh = open(train)
+	probs=ngram.probabilities(ngram.good_turing(ngram.ngram(nmax,filters.unk(filters.shakespeare(fh)))))
 	sentence_generation(train,out,nmax,reps,probs)
 
 def main():

### 2011-02-28 experiment.py
putting figures in paper

@@ -29,7 +29,7 @@ def sentence_generation(train,filename,nmax,reps,probs):
 				pass
 
 def run(train,nmax,out):
-	reps=30
+	reps=2
 	#train='Shakespeare/short.txt'
 	#TODO(tom) This isn't working. I'll fix it later.
 	probs=ngram.probabilities(ngram.good_turing(ngram.ngram(nmax,filters.unk(filters.shakespeare(train)))))

### 2011-02-28 paper/analysis.R

@@ -1,22 +1,24 @@
 get_results=function(file) {read.csv(file,sep='|',colClasses=c('integer','logical','numeric','numeric','character'))}
 
-#results.shakespeare=get_results('Shakespeare results.csv')
-results.tolstoy=get_results('War and Peace results.csv')
+#results.shakespeare=get_results('../Shakespeare results.csv')
+results.tolstoy=get_results('../War and Peace results.csv')
 
-foo=results.tolstoy
-foo.ag=aggregate(foo$perplexity,list(use.fractions=foo$use.fractions,n=foo$n),mean)
+#foo=results.tolstoy
 
 
 #par(mfrow=c(1,2));for (i in c(0,1)){plot((perplexity)~n,data=subset(foo,use.fractions==i),main=paste('use fractions =',i))}
 
-plot(perplexity~n,data=foo,type='n',axes=F,
-	main='Perplexity of generated sentences as a function of n-gram length and probability computation method',
-	xlab='n-gram length',
-	ylab='Perplexity'
-)
-axis(1,at=1:max(foo$n))
-axis(2)
-for (i in c(0,1)){
-	points(perplexity~n,data=subset(foo,use.fractions==i),pch=21,col=NA,bg=1+i)
-	lines(x~n,data=subset(foo.ag,use.fractions==i),col=1+i)
+pp_gen.plot=function(foo){
+	foo.ag=aggregate(foo$perplexity,list(use.fractions=foo$use.fractions,n=foo$n),mean)
+	plot(perplexity~n,data=foo,type='n',axes=F,
+		main='Perplexity of generated sentences as a function of n-gram length and probability computation method',
+		xlab='n-gram length',
+		ylab='Perplexity'
+	)
+	axis(1,at=1:max(foo$n))
+	axis(2)
+	for (i in c(0,1)){
+		points(perplexity~n,data=subset(foo,use.fractions==i),pch=21,col=NA,bg=1+i)
+		lines(x~n,data=subset(foo.ag,use.fractions==i),col=1+i)
+	}
 }

### 2011-02-28 paper/analysis.R
plots in paper

@@ -4,8 +4,6 @@ get_results=function(file) {read.csv(file,sep='|',colClasses=c('integer','logica
 results.tolstoy=get_results('../War and Peace results.csv')
 
 #foo=results.tolstoy
-
-
 #par(mfrow=c(1,2));for (i in c(0,1)){plot((perplexity)~n,data=subset(foo,use.fractions==i),main=paste('use fractions =',i))}
 
 pp_gen.plot=function(foo){

### 2011-02-28 experiment.py
note about p type

@@ -7,11 +7,9 @@ import random
 ngram.good_turing =lambda x:x
 
 def trial(train,n,p,probs):
-	"""n is n-gram size. p is whether it's log, arbitrary precision or float"""
-	#Start timer
+	"""n is n-gram size. p is whether it's log or arbitrary precision"""
 	sentence=ngram.make_sentence(probs[0:(n+1)])
 	pp=ngram.perplexity(probs[0:(n+1)],sentence)
-	#End timer
 	time='NA'
 	return [n,p,pp,time,' '.join(sentence)]
 

### 2011-02-28 experiment.py
move readme

@@ -8,6 +8,7 @@ ngram.good_turing =lambda x:x
 
 def trial(train,n,p,probs):
 	"""n is n-gram size. p is whether it's log or arbitrary precision"""
+	ngram.set_fractions(randint(0,1))
 	sentence=ngram.make_sentence(probs[0:(n+1)])
 	pp=ngram.perplexity(probs[0:(n+1)],sentence)
 	time='NA'
@@ -20,9 +21,9 @@ def sentence_generation(train,filename,nmax,reps,probs):
 	for i in range(0,reps):
 		for n in range(1,nmax+1):
 			try:
-				ngram.set_fractions(True)
-				out.writerow(trial(train,n,True,probs))
-				ngram.set_fractions(False)
+				#ngram.set_fractions(True)
+				#out.writerow(trial(train,n,True,probs))
+				#ngram.set_fractions(False)
 				out.writerow(trial(train,n,False,probs))
 				out_fh.flush()
 			except(AttributeError):

### 2011-02-28 ngram.py

@@ -233,5 +233,3 @@ def main():
 
 if __name__ == "__main__":
 	main()
-
-#ngram.make_sentence(ngram.probabilities(ngram.good_turing(ngram.ngram(3,filters.unk(filters.shakespeare('Shakespeare/short.txt'))))))

### 2011-02-28 ngram.py
don't unk the test set

@@ -225,7 +225,8 @@ def main():
 
 	if args.perplexity is not None:
 		print "this might take a while..."
-		perplex_data = filters.unk(filters.shakespeare(args.perplexity))
+		#perplex_data = filters.unk(filters.shakespeare(args.perplexity))
+		perplex_data = (filters.shakespeare(args.perplexity))
 		print "Perplexity: %s" % perplexity(unked_probs, perplex_data)
 
 	if args.make_sentence:

### 2011-02-28 experiment.py
fixed the experiment
I needed to generate separate probability
dictionaries for arbitrary precision
and logarithms

@@ -8,40 +8,47 @@ ngram.good_turing =lambda x:x
 
 def trial(train,n,p,probs):
 	"""n is n-gram size. p is whether it's log or arbitrary precision"""
-	ngram.set_fractions(randint(0,1))
+	ngram.set_fractions(p)
 	sentence=ngram.make_sentence(probs[0:(n+1)])
 	pp=ngram.perplexity(probs[0:(n+1)],sentence)
 	time='NA'
 	return [n,p,pp,time,' '.join(sentence)]
 
-def sentence_generation(train,filename,nmax,reps,probs):
+def sentence_generation(train,filename,nmax,reps,probs_ap,probs_log):
+	#Open results file
 	out_fh = open(filename, 'wb')
 	out=csv.writer(out_fh, delimiter='|', quotechar='&', quoting=csv.QUOTE_NONE)
 	out.writerow(['n','use fractions','perplexity','time','sentence'])
+	
+	#Write perplexities and other output
 	for i in range(0,reps):
 		for n in range(1,nmax+1):
 			try:
-				#ngram.set_fractions(True)
-				#out.writerow(trial(train,n,True,probs))
-				#ngram.set_fractions(False)
-				out.writerow(trial(train,n,False,probs))
+				out.writerow(trial(train,n,True,probs_ap))
+				out.writerow(trial(train,n,False,probs_log))
 				out_fh.flush()
 			except(AttributeError):
 				pass
 
-def run(train,nmax,out):
-	reps=2
-	#train='Shakespeare/short.txt'
-	#TODO(tom) This isn't working. I'll fix it later.
+def run(train,nmax,reps,out):
+	#Get probabilities
 	fh = open(train)
-	probs=ngram.probabilities(ngram.good_turing(ngram.ngram(nmax,filters.unk(filters.shakespeare(fh)))))
-	sentence_generation(train,out,nmax,reps,probs)
+	ngram.set_fractions(True)
+	probs_ap=ngram.probabilities(ngram.good_turing(ngram.ngram(nmax,filters.unk(filters.shakespeare(fh)))))
+	
+	fh = open(train)
+	ngram.set_fractions(False)
+	probs_log=ngram.probabilities(ngram.good_turing(ngram.ngram(nmax,filters.unk(filters.shakespeare(fh)))))
+	
+	sentence_generation(train,out,nmax,reps,probs_ap,probs_log)
 
 def main():
[…]

### 2011-02-28 ngram.py

@@ -115,10 +115,15 @@ def perplexity(probs, words):
 					break
 				else:
 					temp_words.pop(0)
+	print "prob= %s" % prob
+	print "len(words)= %s" % len(words)
 	if use_fractions:
+		print "Computing perplexity with arbitrary precision"
 		return mp.power((mp.mpf(prob.denom())/mp.mpf(prob.numer())), 1.0/len(words))
 	else:
+		print "Computing perplexity with logarithms"
 		return math.exp(-prob/len(words))
+	#This len(words) may be important for the weird issue we're having
 
 def probabilities(ngrams):
 	global use_fractions

### 2011-02-28 experiment.py
uncomment experiment stuffs

@@ -31,24 +31,24 @@ def sentence_generation(train,filename,nmax,reps,probs_ap,probs_log):
 				pass
 
 def run(train,nmax,reps,out):
-	#Get probabilities
+	#Get probabilities with arbitrary precision
 	fh = open(train)
 	ngram.set_fractions(True)
 	probs_ap=ngram.probabilities(ngram.good_turing(ngram.ngram(nmax,filters.unk(filters.shakespeare(fh)))))
 	
+	#Get probabilities with logs
 	fh = open(train)
 	ngram.set_fractions(False)
 	probs_log=ngram.probabilities(ngram.good_turing(ngram.ngram(nmax,filters.unk(filters.shakespeare(fh)))))
 	
+	#Make sentences
 	sentence_generation(train,out,nmax,reps,probs_ap,probs_log)
 
 def main():
-	nmax=2
-	reps=2
+	nmax=3
+	reps=3
 	run('War and Peace/short.txt',nmax,reps,'War and Peace results.csv')
-	#nmax=5
-	#run('War and Peace/Train.txt',nmax,'War and Peace results.csv')
-	#run('Shakespeare/Train.txt',nmax,'Shakespeare results.csv')
+	run('Shakespeare/Train.txt',nmax,reps,'Shakespeare results.csv')
 
 if __name__ == '__main__':
 	main()

### 2011-02-28 experiment.py
change experiment input slightly

@@ -47,7 +47,7 @@ def run(train,nmax,reps,out):
 def main():
 	nmax=3
 	reps=3
-	run('War and Peace/short.txt',nmax,reps,'War and Peace results.csv')
+	run('War and Peace/Train.txt',nmax,reps,'War and Peace results.csv')
 	run('Shakespeare/Train.txt',nmax,reps,'Shakespeare results.csv')
 
 if __name__ == '__main__':

### 2011-02-28 experiment.py
import data better to R something about experiment

@@ -48,7 +48,7 @@ def main():
 	nmax=3
 	reps=3
 	run('War and Peace/short.txt',nmax,reps,'War and Peace results.csv')
-	run('Shakespeare/Train.txt',nmax,reps,'Shakespeare results.csv')
+	run('Shakespeare/short.txt',nmax,reps,'Shakespeare results.csv')
 
 if __name__ == '__main__':
 	main()

### 2011-02-28 paper/analysis.R

@@ -1,6 +1,11 @@
-get_results=function(file) {read.csv(file,sep='|',colClasses=c('integer','logical','numeric','numeric','character'))}
+get_results=function(file) {
+	read.csv(
+		file,sep='|',header=T,quote='&',
+		colClasses=c('integer','logical','numeric','numeric','character')
+		)
+}
 
-#results.shakespeare=get_results('../Shakespeare results.csv')
+results.shakespeare=get_results('../Shakespeare results.csv')
 results.tolstoy=get_results('../War and Peace results.csv')
 
 #foo=results.tolstoy

### 2011-02-28 experiment.py
change experiment paramaters

@@ -45,8 +45,8 @@ def run(train,nmax,reps,out):
 	sentence_generation(train,out,nmax,reps,probs_ap,probs_log)
 
 def main():
-	nmax=3
-	reps=3
+	nmax=5
+	reps=30
 	#run('War and Peace/short.txt',nmax,reps,'War and Peace results.csv')
 	#run('Shakespeare/short.txt',nmax,reps,'Shakespeare results.csv')
 	run('War and Peace/Train.txt',nmax,reps,'War and Peace results.csv')

### 2011-02-28 experiment_test_corpus.sh
experiment to run lots of perplexities

@@ -1,0 +1,11 @@
+#!/bin/bash
+for n in 1 2 3 4 5 6
+	do for corpus in 'Shakespeare' 'War and Peace'
+		#Header for csv file
+		do echo 'corpus, n-gram length, smoothing, prob, perplexity'
+		echo corpus ', ' $n ', ' 'on'  ', ' 'log' ', ' `./ngram.py -n $n -t $corpus/Train.txt -p $corpus/Test.txt -sl|grep 'Perplexity:'|cut -d ' ' -f 2`
+		echo corpus ', ' $n ', ' 'off' ', ' 'log' ', ' `./ngram.py -n $n -t $corpus/Train.txt -p $corpus/Test.txt  -l|grep 'Perplexity:'|cut -d ' ' -f 2`
+		echo corpus ', ' $n ', ' 'on'  ', ' 'ap'  ', ' `./ngram.py -n $n -t $corpus/Train.txt -p $corpus/Test.txt -s |grep 'Perplexity:'|cut -d ' ' -f 2`
+		echo corpus ', ' $n ', ' 'off' ', ' 'ap'  ', ' `./ngram.py -n $n -t $corpus/Train.txt -p $corpus/Test.txt    |grep 'Perplexity:'|cut -d ' ' -f 2`
+	done
+done

### 2011-02-28 paper/analysis.R

@@ -11,17 +11,27 @@ results.tolstoy=get_results('../War and Peace results.csv')
 #foo=results.tolstoy
 #par(mfrow=c(1,2));for (i in c(0,1)){plot((perplexity)~n,data=subset(foo,use.fractions==i),main=paste('use fractions =',i))}
 
-pp_gen.plot=function(foo){
+pp_gen.plot=function(foo,corpusname){
 	foo.ag=aggregate(foo$perplexity,list(use.fractions=foo$use.fractions,n=foo$n),mean)
 	plot(perplexity~n,data=foo,type='n',axes=F,
-		main='Perplexity of generated sentences as a function of n-gram length and probability computation method',
+		main=paste('Perplexity of generated sentences:',corpusname),
 		xlab='n-gram length',
 		ylab='Perplexity'
+	)
+	legend('topright',
+		c('Arbitrary precision arithmetic','Log probability'),
+		pch=21,
+		col=c(1,2),
+		title='Probability computation method'
 	)
 	axis(1,at=1:max(foo$n))
 	axis(2)
 	for (i in c(0,1)){
-		points(perplexity~n,data=subset(foo,use.fractions==i),pch=21,col=NA,bg=1+i)
+		points(perplexity~n,data=subset(foo,use.fractions==i),
+			pch=21,
+			#col=NA,bg=1+i
+			col=1+i
+		)
 		lines(x~n,data=subset(foo.ag,use.fractions==i),col=1+i)
 	}
 }

### 2011-02-28 experiment_test_corpus.sh
experiment

@@ -2,10 +2,29 @@
 for n in 1 2 3 4 5 6
 	do for corpus in 'Shakespeare' 'War and Peace'
 		#Header for csv file
-		do echo 'corpus, n-gram length, smoothing, prob, perplexity'
-		echo corpus ', ' $n ', ' 'on'  ', ' 'log' ', ' `./ngram.py -n $n -t $corpus/Train.txt -p $corpus/Test.txt -sl|grep 'Perplexity:'|cut -d ' ' -f 2`
-		echo corpus ', ' $n ', ' 'off' ', ' 'log' ', ' `./ngram.py -n $n -t $corpus/Train.txt -p $corpus/Test.txt  -l|grep 'Perplexity:'|cut -d ' ' -f 2`
-		echo corpus ', ' $n ', ' 'on'  ', ' 'ap'  ', ' `./ngram.py -n $n -t $corpus/Train.txt -p $corpus/Test.txt -s |grep 'Perplexity:'|cut -d ' ' -f 2`
-		echo corpus ', ' $n ', ' 'off' ', ' 'ap'  ', ' `./ngram.py -n $n -t $corpus/Train.txt -p $corpus/Test.txt    |grep 'Perplexity:'|cut -d ' ' -f 2`
+		do echo 'corpus, n-gram length, smoothing, prob, perplexity,sentence'
+		echo corpus ', ' $n ', ' 'on'  ', ' 'log' ', ' \
+		#Perplexity
+		`./ngram.py -n $n -t $corpus/Train.txt -sl -p $corpus/Test.txt|grep 'Perplexity:'|cut -d ' ' -f 2` ', ' \
+		#Sentence
+		`./ngram.py -n $n -t $corpus/Train.txt -sl|tail -n 1`
+		
+		echo corpus ', ' $n ', ' 'off' ', ' 'log' ', ' \
+		#Perplexity
+		`./ngram.py -n $n -t $corpus/Train.txt  -l -p $corpus/Test.txt|grep 'Perplexity:'|cut -d ' ' -f 2` ', ' \
+		#Sentence
+		`./ngram.py -n $n -t $corpus/Train.txt  -l|tail -n 1`
+		
+		echo corpus ', ' $n ', ' 'on'  ', ' 'ap'  ', ' \
+		#Perplexity
[…]

### 2011-02-28 paper/paper.Rnw
plots in paper
more paper writing

--- 2011-02-28 paper/paper.tex
@@
 
+<<echo=F>>=
+source('analysis.R')
+@
+
 \begin{document}
+
 \maketitle
+\section*{Implementation}
+
+	Our design is built around dictionaries for each n-gram we consider.  We
+	break the input text into words and add sentence separators (although we do
+	not break the sentences apart, so trigrams or higher can cross sentence
+	boundaries), and then process the list of words, tallying the number of each
+	$n$-gram seen in the dictionary.
+
+	We also can optionally account for unknown words by replacing the first
+	instance of each word with an unknown symbol; this is simply an additional
+	processing step applied to the raw list of words.
+
+	We compute all $n$-gram count dictionaries up to the desired $n$-gram
+	length, and use these to compute the probabilities of each $n$-gram.  This
+	dictionary of probabilities is used to construct sentences and to measure
+	perplexity.  We either use logarithmic or arbitrary-precision arithmetic
+	depending on what the user specifies.
+
+	We implemented Good-Turing smoothing, which adjusts the $n$-gram counts
+	before they are turned into probabilities.  Like unknown-handling, this is
+	simply a pipeline processing step.
+
+	Sentence generation begins with a beginning of sentence symbol, and then
+	chooses words from the $n$-gram dictionary corresponding to the minimum of
+	the desired $n$-gram length or the number of words we have already created.
+
+	Perplexity works similarly, scanning the input list and computing the
+	running sum or product probability, depending on whether we are using
+	arbitrary precision or logarithmic arithmetic.  Because computing the
+	perplexity for each word is simply a matter of looking up a tuple in the
+	appropriate dictionary (possibly containing an unknown), this step is fairly
+	rapid.  If we do not have the appropriate $n$-gram in the dictionary, we
+	first try replacing the last word with an unknown, and, if that fails, we
+	shorten the list of words until a match is found, which, at shortest, is a
+	single unknown by itself.
+
+	\\
+
+	The dictionary approach was simple and intuitive, but fairly slow; the
+	generation of sentences spends the majority of its time on dictionary
+	operations, and uses only moderate amounts of memory.  Much of the time in
+	sentence generation is spent computing the list of $n$-gram options to
+	choose from, since this requires, in our implementation, scanning through
+	the entire $n$-gram dictionary to find matches for the $(n-1)$-gram.  If we
+	precomputed this as we populated the dictionary, substantial time could be
+	saved, although choosing a random element from that list still takes $O(n)$
+	time.
+
 \section*{Sentence Generation}
@@
 
-			The Greeks upon advice , Hath alter'd that good wisdom Whereof I know
-			by their christen names , Unless thou tell her so .
+			The Greeks upon advice , Hath alter'd that good wisdom Whereof I
+			know by their christen names , Unless thou tell her so .
 
@@
 		phenomenon, and is due to the fact that it's quite possible that the
-		preceding $n-1$-gram appears only once in the entire corpus, so the
+		preceding $(n-1)$-gram appears only once in the entire corpus, so the
 		sentence generator has no choice but to choose the next word in the
@@
 
+\section*{Results}
+
+	\subsection*{Pelplexity of test set}
+
+	\subsection*{Perplexity of generated sentences}
+
+\begin{figure}
+\begin{center}
+<<echo=F,fig=T>>=
+pp_gen.plot(results.tolstoy)
+@
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+<<echo=F,fig=T>>=
+pp_gen.plot(results.shakespeare)
+@
+\end{center}
+\end{figure}
+
 \end{document}

### 2011-02-28 paper/paper.Rnw

@@
 \begin{center}
-<<echo=F,fig=T>>=
-pp_gen.plot(results.tolstoy)
+<<echo=F,fig=T,height=9>>=
+par(mfrow=c(2,1))
+pp_gen.plot(results.shakespeare,'Shakespeare')
+pp_gen.plot(results.tolstoy,'War and Peace')
 @
@@
 <<echo=F,fig=T>>=
-pp_gen.plot(results.shakespeare)
+#pp_gen.plot(results.shakespeare,'Shakespeare')
 @

### 2011-02-28 experiment_test_corpus.sh
ignore paper.aux

@@ -1,30 +1,18 @@
 #!/bin/bash
-for n in 1 2 3 4 5 6
-	do for corpus in 'Shakespeare' 'War and Peace'
+#TRAIN='Train.txt'
+#TEST='Test.txt'
+TRAIN='shorter.txt'
+TEST='short.txt'
+echo 'corpus, n-gram length, smoothing, prob, perplexity,sentence'
+for n in 1 #2 #3 4 5 6
+	do for corpus in 'Shakespeare' #'War and Peace'
 		#Header for csv file
-		do echo 'corpus, n-gram length, smoothing, prob, perplexity,sentence'
-		echo corpus ', ' $n ', ' 'on'  ', ' 'log' ', ' \
-		#Perplexity
-		`./ngram.py -n $n -t $corpus/Train.txt -sl -p $corpus/Test.txt|grep 'Perplexity:'|cut -d ' ' -f 2` ', ' \
-		#Sentence
-		`./ngram.py -n $n -t $corpus/Train.txt -sl|tail -n 1`
+		do echo corpus ', ' $n ', ' 'on'  ', ' 'log' ', ' `./ngram.py -n $n -t $corpus/$TRAIN -sl -p $corpus/$TEST|grep '^Perplexity:'|cut -d ' ' -f 2` ', ' `./ngram.py -n $n -t $corpus/$TRAIN -slm|tail -n 1`
 		
-		echo corpus ', ' $n ', ' 'off' ', ' 'log' ', ' \
-		#Perplexity
-		`./ngram.py -n $n -t $corpus/Train.txt  -l -p $corpus/Test.txt|grep 'Perplexity:'|cut -d ' ' -f 2` ', ' \
-		#Sentence
-		`./ngram.py -n $n -t $corpus/Train.txt  -l|tail -n 1`
+		echo    corpus ', ' $n ', ' 'off' ', ' 'log' ', ' `./ngram.py -n $n -t $corpus/$TRAIN  -l -p $corpus/$TEST|grep '^Perplexity:'|cut -d ' ' -f 2` ', ' `./ngram.py -n $n -t $corpus/$TRAIN  -lm|tail -n 1`
 		
-		echo corpus ', ' $n ', ' 'on'  ', ' 'ap'  ', ' \
-		#Perplexity
[…]

### 2011-02-28 paper/paper.Rnw
"problematic"

@@
 
-	Our design is built around dictionaries for each n-gram we consider.  We
+	Our design is built around dictionaries for each $n$-gram we consider.  We
 	break the input text into words and add sentence separators (although we do
@@
 	from War and Peace.  These corpora were chosen because they were determined
-	at rough glance to be the easiest to parse and had the fewest problematic.
+	at rough glance to be the easiest to parse. %and had the fewest problematic.
 
@@
 
-\begin{figure}
-\begin{center}
-<<echo=F,fig=T>>=
-#pp_gen.plot(results.shakespeare,'Shakespeare')
-@
-\end{center}
-\end{figure}
-
 \end{document}

### 2011-02-28 experiment_test_corpus.sh
perplexity experiment should work

@@ -1,11 +1,19 @@
 #!/bin/bash
-#TRAIN='Train.txt'
-#TEST='Test.txt'
-TRAIN='shorter.txt'
-TEST='short.txt'
+#Calculate perplexities and
+#generate sentences with different
+	#different corpora
+	#different n-gram lengths
+	#smoothing on or off
+	#different probability handling
+
+TRAIN='Train.txt'
+TEST='Test.txt'
+#Note: This fails on really really small corpora.
+#TRAIN='shorter.txt'
+#TEST='short.txt'
 echo 'corpus, n-gram length, smoothing, prob, perplexity,sentence'
-for n in 1 #2 #3 4 5 6
-	do for corpus in 'Shakespeare' #'War and Peace'
+for n in 1 2 3 4 5 6
+	do for corpus in 'Shakespeare' 'War_and_Peace'
 		#Header for csv file
 		do echo corpus ', ' $n ', ' 'on'  ', ' 'log' ', ' `./ngram.py -n $n -t $corpus/$TRAIN -sl -p $corpus/$TEST|grep '^Perplexity:'|cut -d ' ' -f 2` ', ' `./ngram.py -n $n -t $corpus/$TRAIN -slm|tail -n 1`
 		

### 2011-02-28 experiment_test_corpus.sh
skip arbitrary precision for now

@@ -6,21 +6,26 @@
 	#smoothing on or off
 	#different probability handling
 
-TRAIN='Train.txt'
-TEST='Test.txt'
+#TRAIN='Train.txt'
+#TEST='Test.txt'
+TRAIN='short1.txt'
+TEST='short2.txt'
+
 #Note: This fails on really really small corpora.
 #TRAIN='shorter.txt'
 #TEST='short.txt'
+
 echo 'corpus, n-gram length, smoothing, prob, perplexity,sentence'
 for n in 1 2 3 4 5 6
 	do for corpus in 'Shakespeare' 'War_and_Peace'
 		#Header for csv file
-		do echo corpus ', ' $n ', ' 'on'  ', ' 'log' ', ' `./ngram.py -n $n -t $corpus/$TRAIN -sl -p $corpus/$TEST|grep '^Perplexity:'|cut -d ' ' -f 2` ', ' `./ngram.py -n $n -t $corpus/$TRAIN -slm|tail -n 1`
+		do echo $corpus ', ' $n ', ' 'on'  ', ' 'log' ', ' `./ngram.py -n $n -t $corpus/$TRAIN -sl -p $corpus/$TEST|grep '^Perplexity:'|cut -d ' ' -f 2` ', ' `./ngram.py -n $n -t $corpus/$TRAIN -slm|tail -n 1`
 		
-		echo    corpus ', ' $n ', ' 'off' ', ' 'log' ', ' `./ngram.py -n $n -t $corpus/$TRAIN  -l -p $corpus/$TEST|grep '^Perplexity:'|cut -d ' ' -f 2` ', ' `./ngram.py -n $n -t $corpus/$TRAIN  -lm|tail -n 1`
+		echo    $corpus ', ' $n ', ' 'off' ', ' 'log' ', ' `./ngram.py -n $n -t $corpus/$TRAIN  -l -p $corpus/$TEST|grep '^Perplexity:'|cut -d ' ' -f 2` ', ' `./ngram.py -n $n -t $corpus/$TRAIN  -lm|tail -n 1`
+
+#Skip ap for now		
[…]

### 2011-02-28 experiment_test_corpus.sh
set train and test

@@ -6,10 +6,10 @@
 	#smoothing on or off
 	#different probability handling
 
-#TRAIN='Train.txt'
-#TEST='Test.txt'
-TRAIN='short1.txt'
-TEST='short2.txt'
+TRAIN='Train.txt'
+TEST='Test.txt'
+#TRAIN='short1.txt'
+#TEST='short2.txt'
 
 #Note: This fails on really really small corpora.
 #TRAIN='shorter.txt'

### 2011-03-01 paper/analysis.R
colors and remove unigrams from plot

@@ -5,6 +5,8 @@ get_results=function(file) {
 		)
 }
 
+mycols=c('00000060','FFFF0060')
+
 results.shakespeare=get_results('../Shakespeare results.csv')
 results.tolstoy=get_results('../War and Peace results.csv')
 
@@ -21,7 +23,7 @@ pp_gen.plot=function(foo,corpusname){
 	legend('topright',
 		c('Arbitrary precision arithmetic','Log probability'),
 		pch=21,
-		col=c(1,2),
+		col=mycols[c(1,2)],
 		title='Probability computation method'
 	)
 	axis(1,at=1:max(foo$n))
@@ -30,8 +32,8 @@ pp_gen.plot=function(foo,corpusname){
 		points(perplexity~n,data=subset(foo,use.fractions==i),
 			pch=21,
 			#col=NA,bg=1+i
-			col=1+i
+			col=mycols[1+i]
 		)
-		lines(x~n,data=subset(foo.ag,use.fractions==i),col=1+i)
+		lines(x~n,data=subset(foo.ag,use.fractions==i),col=mycols[1+i])
 	}
 }

### 2011-03-01 paper/analysis.R
colors

@@ -6,7 +6,7 @@ get_results=function(file) {
 }
 
 results.shakespeare=get_results('../Shakespeare results.csv')
-results.tolstoy=get_results('../War and Peace results.csv')
+results.tolstoy=get_results('war_and_peace.csv')
 
 #foo=results.tolstoy
 #par(mfrow=c(1,2));for (i in c(0,1)){plot((perplexity)~n,data=subset(foo,use.fractions==i),main=paste('use fractions =',i))}

### 2011-03-01 paper/analysis.R
hash mark in color

@@ -5,7 +5,7 @@ get_results=function(file) {
 		)
 }
 
-mycols=c('00000060','FFFF0060')
+mycols=c('#00000060','#FFFF0060')
 
 results.shakespeare=get_results('../Shakespeare results.csv')
 results.tolstoy=get_results('../War and Peace results.csv')

### 2011-03-01 paper/paper.Rnw
minor capitalization
perplexity experiment should work
colors and remove unigrams from plot
colors
paper

@@
 \newcommand{\osn}{\oldstylenums}
-\title{Arbitrary Precision and Arbitrary $n$-grams}
+\title{Arbitrary Precision and Arbitrary $n$-Grams}
 \author{Alec Story and Thomas Levine\\\small{avs\osn{38} and tkl\osn{22}}}
+
+\newcommand\ngram{$n$-gram}
+\newcommand\loga{logarithmic arithmetic}
 
@@
 
-	\\
+	%\\
 
@@
 	not convert the first occurrence of each word to an unknown token.
+
+	%We trained our system on the \emph{Shakespeare} and \emph{War and Peace} corpora.
 
@@
 
+\section*{Experimental design}
+We trained our model on the training data and then
+ran two experiments to study the effect of \ngram length and
+probability computation type on perplexity on the two corpora.
+The two probability computation types were arbitrary precision
+arithmetic and \loga.
+For each experiment, we tested \ngram length of $n$=$1$ to $5$.
+
+We hypothesized that longer $n$-grams would yield lower perplexity
+and that the two probability computation types would yield similar perplexity.
+
+\subsection*{Perplexity of test set}
+We evaluated our model on the test set by computing
+the perplexity of the test set. In addition to the variables
+above, we ran this test with smoothing on and with smoothing off.
+We hypothesized that smoothing would reduce perplexity.
+
+\subsection*{Perplexity of generated sentences}
+We also evaluated our model on sentences generated by our model.
+We did this because sentence generation involves substantial
+probability computations, where the difference in probability
+computation type would be relevant.
+
+For each combination of the intependent variables, we
+generated 23 sentences from the \emph{War and Peace} corpus
+and tested their perplexity. %ADD SHAKESPEARE
+
 \section*{Results}
@@
 \subsection*{Perplexity of generated sentences}
+Perplexity of generated sentences decreased with increased
+\ngram length
 
@@
 \end{center}
+\caption{\label{fig:gen}Perplexities of generated sentences by $n$-gram length and probability computation method for the two corpora.}
 \end{figure}

### 2011-03-01 paper/analysis.R

@@ -5,10 +5,10 @@ get_results=function(file) {
 		)
 }
 
-mycols=c('#00000060','#FFFF0060')
+mycols=rev(c('#00000060','#FF000060'))
 
 results.shakespeare=get_results('../Shakespeare results.csv')
-results.tolstoy=get_results('../War and Peace results.csv')
+results.tolstoy=get_results('war_and_peace.csv')
 
 #foo=results.tolstoy
 #par(mfrow=c(1,2));for (i in c(0,1)){plot((perplexity)~n,data=subset(foo,use.fractions==i),main=paste('use fractions =',i))}
@@ -23,7 +23,7 @@ pp_gen.plot=function(foo,corpusname){
 	legend('topright',
 		c('Arbitrary precision arithmetic','Log probability'),
 		pch=21,
-		col=mycols[c(1,2)],
+		col=c(2,1),
 		title='Probability computation method'
 	)
 	axis(1,at=1:max(foo$n))

### 2011-03-01 paper/analysis.R
test pp data

@@ -37,3 +37,6 @@ pp_gen.plot=function(foo,corpusname){
 		lines(x~n,data=subset(foo.ag,use.fractions==i),col=mycols[1+i])
 	}
 }
+
+
+read.csv('test_pp_log.csv')

### 2011-03-01 paper/paper.Rnw
test set pp
paper changes

--- 2011-02-28 paper/paper.Rnw
@@
 \newcommand{\osn}{\oldstylenums}
-\title{Arbitrary Precision and Arbitrary $n$-grams}
+\title{Arbitrary Precision and Arbitrary $n$-Grams}
 \author{Alec Story and Thomas Levine\\\small{avs\osn{38} and tkl\osn{22}}}
@@
 	from War and Peace.  These corpora were chosen because they were determined
-	at rough glance to be the easiest to parse. %and had the fewest problematic.
+	at rough glance to be the easiest to parse.
 
@@
 		implemented it using logarithmic arithmetic).  This can lead to slower
-		processing, but in our profiling, most of the time was spent in
-		dictionary manipulation rather than mathematics.  This approach
-		guarantees that there is no data loss until we move into processing
-		floats during exponentiation.
+		processing, but	guarantees that there is no data loss until we move into
+		processing floats during exponentiation.
 
@@
 		Our hypothesis was that the arbitrary precision math would reveal a
-		small amount of error in computing the perplexity logarithmically.
+		small amount of error in computing the perplexity logarithmically.  Our
+		results seem to indicate that any such error is very small (differing
+		only in the last digit in measuring the perplexity of whole corpora),
+		but that arbitrary-precision mathematics is much slower than floating
+		point mathematics with logs when finding the product of many numbers.
+		See \ref{runtimes}:
+
+		\begin{table}
+		\caption{Running times with arbitrary and floating-point precision in
+		seconds.  Average of 10 trials, trained on the Shakespeare training
+		corpus, perplexity calculated for Shakespeare training corpus}
+		\label{runtimes}
+		\begin{tabular}{l | r r}
+			Operation & Arbitrary Precision & Logarithmic \\
+			\hline
+			Trigram tallying & 136.936949968 & 137.023640156\\
+			Probability computation & 63.5648961067 & 60.1038038731 \\
+			Perplexity computation & \textbf{990.322752953} & 65.2168021202 \\
+		\end{tabular}
+		\end{table}
 
@@
 
+\begin{figure}
+\begin{center}
+<<echo=F,fig=T,height=9>>=
+par(mfrow=c(2,1))
+pp_gen.plot(subset(results.shakespeare,n>1),'Shakespeare')
+pp_gen.plot(subset(results.tolstoy,n>1),'War and Peace')
+@
+\end{center}
+\caption{Perplexities for n-grams of length greater than $n$=1}
+\end{figure}
+
 \end{document}

### 2011-03-01 paper/paper.Rnw

--- 2011-03-01 paper/paper.Rnw
@@
 
-	%We trained our system on the \emph{Shakespeare} and \emph{War and Peace} corpora.
-
 	Each subsection contains two examples from the Shakespeare corpus and two
 	from War and Peace.  These corpora were chosen because they were determined
-	at rough glance to be the easiest to parse. %and had the fewest problematic.
+	at rough glance to be the easiest to parse.
 

### 2011-03-01 paper/analysis.R
shakespeare

@@ -7,7 +7,7 @@ get_results=function(file) {
 
 mycols=c('#00000060','#FFFF0060')
 
-results.shakespeare=get_results('../Shakespeare results.csv')
+results.shakespeare=get_results('../shakespeare.csv')
 results.tolstoy=get_results('war_and_peace.csv')
 
 #foo=results.tolstoy

### 2011-03-01 paper/paper.Rnw
further paper changes
paper stuff

@@
 	single unknown by itself.
-
-	%\\
 
@@
 		point mathematics with logs when finding the product of many numbers.
-		See \ref{runtimes}:
+		See table \ref{runtimes}.  This suggests that the logarithmic
+		computation is indeed an accurate way to compute perplexity,
+		particularly in light of the more than \osn{10}-fold speed improvements
+		it results in.
 
-		\begin{table}
-		\caption{Running times with arbitrary and floating-point precision in
-		seconds.  Average of 10 trials, trained on the Shakespeare training
-		corpus, perplexity calculated for Shakespeare training corpus}
-		\label{runtimes}
+		\begin{table}[h]
 		\begin{tabular}{l | r r}
@@
 		\end{tabular}
+		\caption{Running times with arbitrary and floating-point precision in
+		seconds.  Average of \osn{10} trials, trained on the Shakespeare
+		training corpus, perplexity calculated for Shakespeare training corpus}
+		\label{runtimes}
 		\end{table}

### 2011-03-01 paper/analysis.R

@@ -7,7 +7,7 @@ get_results=function(file) {
 
 mycols=rev(c('#00000060','#FF000060'))
 
-results.shakespeare=get_results('../shakespeare.csv')
+results.shakespeare=get_results('shakespeare.csv')
 results.tolstoy=get_results('war_and_peace.csv')
 
 #foo=results.tolstoy

### 2011-03-01 timer.py

@@ -1,0 +1,24 @@
+#!/usr/bin/env/python
+import filters
+import ngram
+
+train = open('Shakespeare/Train.txt')
+test = open('Shakespeare/Test.txt')
+
+ngram.set_fractions(True)
+train_words = filters.unk(filters.shakespeare(train))
+ngrams = ngram.ngram(3, train_words)
+probs = ngram.probabilities(ngrams)
+test_words = filters.shakespeare(test)
+
+def test(use_fractions):
+	ngram.set_fractions(use_fractions)
+	print ngram.perplexity(probs, test_words)
+
+if __name__=='__main__':
+	print "testing now"
+	from timeit import Timer
+	t = Timer("test(True)", "from __main__ import test")
+	#u = Timer("test(False)", "from __main__ import test")
+	print "fractions on: %s" % t.timeit(number=10)
+	#print "fractions off: %s" % u.timeit(number=10)

### 2011-03-01 paper/paper.Rnw
test pp

--- 2011-03-01 paper/paper.Rnw
@@
 
+	%We trained our system on the \emph{Shakespeare} and \emph{War and Peace} corpora.
+
 	Each subsection contains two examples from the Shakespeare corpus and two
 	from War and Peace.  These corpora were chosen because they were determined
-	at rough glance to be the easiest to parse.
+	at rough glance to be the easiest to parse. %and had the fewest problematic.
 
@@
 		implemented it using logarithmic arithmetic).  This can lead to slower
-		processing, but	guarantees that there is no data loss until we move into
-		processing floats during exponentiation.
+		processing, but in our profiling, most of the time was spent in
+		dictionary manipulation rather than mathematics.  This approach
+		guarantees that there is no data loss until we move into processing
+		floats during exponentiation.
 
@@
 		Our hypothesis was that the arbitrary precision math would reveal a
-		small amount of error in computing the perplexity logarithmically.  Our
-		results seem to indicate that any such error is very small (differing
-		only in the last digit in measuring the perplexity of whole corpora),
-		but that arbitrary-precision mathematics is much slower than floating
-		point mathematics with logs when finding the product of many numbers.
-		See \ref{runtimes}:
-
-		\begin{table}
-		\caption{Running times with arbitrary and floating-point precision in
-		seconds.  Average of 10 trials, trained on the Shakespeare training
-		corpus, perplexity calculated for Shakespeare training corpus}
-		\label{runtimes}
-		\begin{tabular}{l | r r}
-			Operation & Arbitrary Precision & Logarithmic \\
-			\hline
-			Trigram tallying & 136.936949968 & 137.023640156\\
-			Probability computation & 63.5648961067 & 60.1038038731 \\
-			Perplexity computation & \textbf{990.322752953} & 65.2168021202 \\
-		\end{tabular}
-		\end{table}
+		small amount of error in computing the perplexity logarithmically.
 
@@
 
+\begin{figure}
+\begin{center}
+<<echo=F,fig=T,height=9>>=
+test.plot()
+@
+\end{center}
+\caption{\label{fig:test}Perplexities of the test corpora by $n$-gram length and probability computation method for models trained on the respective training corpora.}
+\end{figure}
+
 \subsection*{Perplexity of generated sentences}

### 2011-03-01 paper/analysis.R

@@ -38,5 +38,29 @@ pp_gen.plot=function(foo,corpusname){
 	}
 }
 
+test.plot=function(){
+	par(mfrow=c(2,1))
+	test=read.csv('test_pp.csv')
+	plot(perplexity~n.gram_length,data=subset(test,corpus=='Shakespeare'),type='n',
+	xlab='n-gram length',ylab='Perplexity',main='Perplexity of the Shakespeare test corpus'
+	)
+	points(perplexity~n.gram_length,data=subset(test,prob=='ap'&smoothing=='on'&corpus=='Shakespeare'),type='l')
+	points(perplexity~n.gram_length,data=subset(test,prob=='ap'&smoothing=='off'&corpus=='Shakespeare'),type='l',col=2)
+	#points(perplexity~n.gram_length,data=subset(test,prob=='log'&smoothing=='off'&corpus=='Shakespeare'),type='l')
+	legend('topright',
+			c('Smoothing on','Smoothing off'),
+			lty=1,
+			col=c(1,2)
+		)
 
-read.csv('test_pp_log.csv')
+	plot(perplexity~n.gram_length,data=subset(test,corpus=='War_and_Peace'),type='n',
+	xlab='n-gram length',ylab='Perplexity',main='Perplexity of the War and Peace test corpus'
+	)
+	points(perplexity~n.gram_length,data=subset(test,prob=='ap'&smoothing=='on'&corpus=='War_and_Peace'),type='l')
+	points(perplexity~n.gram_length,data=subset(test,prob=='ap'&smoothing=='off'&corpus=='War_and_Peace'),type='l',col=2)
+	legend('topright',
+			c('Smoothing on','Smoothing off'),
+			lty=1,
+			col=c(1,2)
+		)
+}		

### 2011-03-01 paper/paper.Rnw
blarg

--- 2011-03-01 paper/paper.Rnw
@@
 above, we ran this test with smoothing on and with smoothing off.
-We hypothesized that smoothing would reduce perplexity.
 
@@
 generated 23 sentences from the \emph{War and Peace} corpus
-and tested their perplexity. %ADD SHAKESPEARE
+and tested their perplexity. We did the same the Shakespeare corpus
+but generated only 7 sentences for each treatment combination.
 
@@
 \subsection*{Pelplexity of test set}
+For each corpus, perplexity of test set decreased when the models
+were run for longer $n$-grams, and the perplexity
+leveled off starting at $n=3$ (figure \ref{fig:test}). Probability computation type had no effect.
+
+Smoothing slightly increased perplexity. The increase
+was larger for larger $n$.
+
+\begin{figure}
+\begin{center}
+<<echo=F,fig=T,height=9>>=
+test.plot()
+@
+\end{center}
+\caption{\label{fig:test}Perplexities of the test corpora by $n$-gram length and probability computation method for models trained on the respective training corpora.}
+\end{figure}
 
 \subsection*{Perplexity of generated sentences}
-Perplexity of generated sentences decreased with increased
-\ngram length
+Perplexity of generated sentences decreased with when longer $n$-grams
+were used in the model (figure \ref{fig:gen}). The perplexity
+leveled off starting at $n=3$. Probability computation type did not
+affect perplexity.
 

### 2011-03-01 paper/paper.Rnw
spellcheck

@@
 
-For each combination of the intependent variables, we
+For each combination of the independent variables, we
 generated 23 sentences from the \emph{War and Peace} corpus
@@
 
-\subsection*{Pelplexity of test set}
+\subsection*{Perplexity of test set}
 For each corpus, perplexity of test set decreased when the models

### 2011-03-01 paper/paper.Rnw
paper changes

--- 2011-03-01 paper/paper.Rnw
@@
 
-		We used python's +gmpy+ package to represent the probabilities as
+		We used python's \verb+gmpy+ package to represent the probabilities as
 		arbitrary rational numbers (python's integers are already
-		arbitrary-precision), and +mpmath+ to perform the final exponentiation
-		to calculate the perplexity (the $n$-th root).  The libraries could only
-		compute non-integer powers of integers or floats, and integer powers
-		were rounded to the nearest integer, and therefore, inaccurate, so
-		(very high precision) floats provided the best results.  This final
-		stage of computation of perplexity was the only point where imprecise
-		mathematics was used.
+		arbitrary-precision), and \verb+mpmath+ to perform the final
+		exponentiation to calculate the perplexity (the $n$-th root).  The
+		libraries could only compute non-integer powers of integers or floats,
+		and integer powers were rounded to the nearest integer, and therefore,
+		inaccurate, so (very high precision) floats provided the best results.
+		This final stage of computation of perplexity was the only point where
+		imprecise mathematics was used. See table \ref{values}
+
+		\begin{table}[p]
+		\begin{tabular}{l r l l | l}
+			Corpus & n-gram length & smoothing & math & perplexity\\
+			\hline
+			Shakespeare &  1 &  on &  log &  499.299251351\\
+			Shakespeare &  1 &  on &  ap &  499.299251353146\\
+			Shakespeare &  1 &  off &  log &  487.213020203\\
+			Shakespeare &  1 &  off &  ap &  487.213020203167\\
+			War and peace &  1 &  on &  log &  439.45958473\\
+			War and peace &  1 &  on &  ap &  439.459584729599\\
+			War and peace &  1 &  off &  log &  431.095976221\\
+			War and peace &  1 &  off &  ap &  431.095976221135\\
+			Shakespeare &  2 &  on &  log &  61.1651128384\\
+			Shakespeare &  2 &  on &  ap &  61.1651128383365\\
+			Shakespeare &  2 &  off &  log &  53.723243187\\
+			Shakespeare &  2 &  off &  ap &  53.7232431870133\\
+			War and peace &  2 &  on &  log &  50.5728265194\\
+			War and peace &  2 &  on &  ap &  50.5728265193686\\
+			War and peace &  2 &  off &  log &  44.4939985115\\
+			War and peace &  2 &  off &  ap &  44.4939985115293\\
+			Shakespeare &  3 &  on &  log &  42.3740593974\\
+			Shakespeare &  3 &  on &  ap &  42.3740593973802\\
+			Shakespeare &  3 &  off &  log &  29.644105241\\
+			Shakespeare &  3 &  off &  ap &  29.644105240972\\
+			War and peace &  3 &  on &  log &  33.5298099459\\
+			War and peace &  3 &  on &  ap &  33.529809945862\\
+			War and peace &  3 &  off &  log &  23.6337013897\\
+			War and peace &  3 &  off &  ap &  23.6337013896575\\
+			Shakespeare &  4 &  on &  log &  41.6859932104\\
+			Shakespeare &  4 &  on &  ap &  41.6859932104407\\
+			Shakespeare &  4 &  off &  log &  25.5850207553\\
+			Shakespeare &  4 &  off &  ap &  25.5850207552466\\
+			War and peace &  4 &  on &  log &  33.0110361869\\
+			War and peace &  4 &  on &  ap &  33.0110361868792\\
+			War and peace &  4 &  off &  log &  20.5857956417\\
+			War and peace &  4 &  off &  ap &  20.5857956417055\\
+			Shakespeare &  5 &  on &  log &  42.4661223625 \\
+			Shakespeare &  5 &  on &  ap &  42.4661223624809\\
+			Shakespeare &  5 &  off &  log &  25.0344976529\\
+			Shakespeare &  5 &  off &  ap &  25.034497652926\\
+			War and peace &  5 &  on &  log &  33.6199337718\\
+			War and peace &  5 &  on &  ap &  33.6199337717408\\
+			War and peace &  5 &  off &  log &  20.0240585504\\
+			War and peace &  5 &  off &  ap &  20.0240585503833\\
+			Shakespeare &  6 &  on &  log &  42.7513356663 \\
+			Shakespeare &  6 &  on &  ap &  42.7513356662635\\
+			Shakespeare &  6 &  off &  log &  24.9268403396 \\
+			Shakespeare &  6 &  off &  ap &  24.9268403395792\\
+		\end{tabular}
+		\caption{Perplexity values for various n-gram parameters}
+		\label{values}
+		\end{table}
 
@@
 above, we ran this test with smoothing on and with smoothing off.
+We hypothesized that smoothing would reduce perplexity.
 
@@
 generated 23 sentences from the \emph{War and Peace} corpus
-and tested their perplexity. We did the same the Shakespeare corpus
-but generated only 7 sentences for each treatment combination.
+and tested their perplexity. %ADD SHAKESPEARE
 
@@
 \subsection*{Pelplexity of test set}
-For each corpus, perplexity of test set decreased when the models
-were run for longer $n$-grams, and the perplexity
-leveled off starting at $n=3$ (figure \ref{fig:test}). Probability computation type had no effect.
-
-Smoothing slightly increased perplexity. The increase
-was larger for larger $n$.
-
-\begin{figure}
-\begin{center}
-<<echo=F,fig=T,height=9>>=
-test.plot()
-@
-\end{center}
-\caption{\label{fig:test}Perplexities of the test corpora by $n$-gram length and probability computation method for models trained on the respective training corpora.}
-\end{figure}
 
 \subsection*{Perplexity of generated sentences}
-Perplexity of generated sentences decreased with when longer $n$-grams
-were used in the model (figure \ref{fig:gen}). The perplexity
-leveled off starting at $n=3$. Probability computation type did not
-affect perplexity.
+Perplexity of generated sentences decreased with increased
+\ngram length
 
@@
 
+\section*{Conclusion}
+
+	Arbitrary precision does not result in more accurate perplexity computations
+	over logarithmic computation, at least to floating point precision on the
+	output of the perplexity, and such precision is probably dwarfed by noise in
+	the inputs.  Such computations, however, take substantially longer to
+	compute, so arbitrary precision computations are probably not worth
+	performing.
+
+	As expected, lengthening n-grams to moderate values decreases perplexity of
+	the test corpus, but increasing the n-gram length beyond 3 did not yield
+	benefits, so it appears that trigrams are reasonable for use computing
+	perplexity.
+
 \end{document}

### 2011-03-01 paper/paper.Rnw

@@
 
-For each combination of the intependent variables, we
+For each combination of the independent variables, we
 generated 23 sentences from the \emph{War and Peace} corpus
-and tested their perplexity. %ADD SHAKESPEARE
+and tested their perplexity. We did the same the Shakespeare corpus
+but generated only 7 sentences for each treatment combination.
 
@@
 
-\subsection*{Pelplexity of test set}
+\subsection*{Perplexity of test set}
+For each corpus, perplexity of test set decreased when the models
+were run for longer $n$-grams, and the perplexity
+leveled off starting at $n=3$ (figure \ref{fig:test}). Probability computation type had no effect.
+
+Smoothing slightly increased perplexity. The increase
+was larger for larger $n$.
+
+\begin{figure}
+\begin{center}
+<<echo=F,fig=T,height=9>>=
+test.plot()
+@
+\end{center}
+\caption{\label{fig:test}Perplexities of the test corpora by $n$-gram length and probability computation method for models trained on the respective training corpora.}
+\end{figure}
 
 \subsection*{Perplexity of generated sentences}
-Perplexity of generated sentences decreased with increased
-\ngram length
+Perplexity of generated sentences decreased with when longer $n$-grams
+were used in the model (figure \ref{fig:gen}). The perplexity
+leveled off starting at $n=3$. Probability computation type did not
+affect perplexity.
 

### 2011-03-01 paper/paper.Rnw
move

@@
 \author{Alec Story and Thomas Levine\\\small{avs\osn{38} and tkl\osn{22}}}
-
-\newcommand\ngram{$n$-gram}
-\newcommand\loga{logarithmic arithmetic}
 
@@
 
-		Our hypothesis was that the arbitrary precision math would reveal a
-		small amount of error in computing the perplexity logarithmically.  Our
-		results seem to indicate that any such error is very small (differing
-		only in the last digit in measuring the perplexity of whole corpora),
-		but that arbitrary-precision mathematics is much slower than floating
-		point mathematics with logs when finding the product of many numbers.
-		See table \ref{runtimes}.  This suggests that the logarithmic
-		computation is indeed an accurate way to compute perplexity,
-		particularly in light of the more than \osn{10}-fold speed improvements
-		it results in.
-
-		\begin{table}[h]
-		\begin{tabular}{l | r r}
-			Operation & Arbitrary Precision & Logarithmic \\
-			\hline
-			Trigram tallying & 136.936949968 & 137.023640156\\
-			Probability computation & 63.5648961067 & 60.1038038731 \\
-			Perplexity computation & \textbf{990.322752953} & 65.2168021202 \\
-		\end{tabular}
-		\caption{Running times with arbitrary and floating-point precision in
-		seconds.  Average of \osn{10} trials, trained on the Shakespeare
-		training corpus, perplexity calculated for Shakespeare training corpus}
-		\label{runtimes}
-		\end{table}
-
 	\section*{Arbitrary $n$-grams}
@@
 We trained our model on the training data and then
-ran two experiments to study the effect of \ngram length and
+ran two experiments to study the effect of $n$-gram length and
 probability computation type on perplexity on the two corpora.
 The two probability computation types were arbitrary precision
-arithmetic and \loga.
-For each experiment, we tested \ngram length of $n$=$1$ to $5$.
+arithmetic and logarithmic arithmetic.
+For each experiment, we tested $n$-gram length of $n$=$1$ to $5$.
 
@@
 
+\subsection*{Arbitrary precision arithmetic}
+		Our hypothesis was that the arbitrary precision math would reveal a
+		small amount of error in computing the perplexity logarithmically.  Our
+		results seem to indicate that any such error is very small (differing
+		only in the last digit in measuring the perplexity of whole corpora),
+		but that arbitrary-precision mathematics is much slower than floating
+		point mathematics with logs when finding the product of many numbers.
+		See table \ref{runtimes}.  This suggests that the logarithmic
+		computation is indeed an accurate way to compute perplexity,
+		particularly in light of the more than \osn{10}-fold speed improvements
+		it results in.
+
+		\begin{table}[h]
+		\begin{tabular}{l | r r}
+			Operation & Arbitrary Precision & Logarithmic \\
+			\hline
+			Trigram tallying & 136.936949968 & 137.023640156\\
+			Probability computation & 63.5648961067 & 60.1038038731 \\
+			Perplexity computation & \textbf{990.322752953} & 65.2168021202 \\
+		\end{tabular}
+		\caption{Running times with arbitrary and floating-point precision in
+		seconds.  Average of \osn{10} trials, trained on the Shakespeare
+		training corpus, perplexity calculated for Shakespeare training corpus}
+		\label{runtimes}
+		\end{table}
+
 \section*{Conclusion}

### 2011-03-01 experiment.py
experiment works better

@@ -49,7 +49,7 @@ def main():
 	reps=30
 	#run('War and Peace/short.txt',nmax,reps,'War and Peace results.csv')
 	#run('Shakespeare/short.txt',nmax,reps,'Shakespeare results.csv')
-	#run('War and Peace/Train.txt',nmax,reps,'War and Peace results.csv')
+	run('War and Peace/Train.txt',nmax,reps,'War and Peace results.csv')
 	run('Shakespeare/Train.txt',nmax,reps,'Shakespeare results.csv')
 
 if __name__ == '__main__':
